Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-10-18 paper.tex
More text

\documentclass{article}

\title{Boutiques: a flexible framework for automated pipeline integration in computing platforms}

@@

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, \\
        Pierre Bellec, Remi Bernard, John Flavin, Krzysztof J. Gorgolewski, \\
        Nathaniel Kofalt, Pierre-Olivier Quirion, Pierre Rioux, \\
        Gunnar Schaeffer, Marc-\'Etienne Rousseau, Alan C. Evans}

@@

@@

@@

@@

@@

@@

\section{Evaluation}

\subsection{Implementation}

@@

@@

>>>>

\documentclass{article}
\usepackage{minted}

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}

\title{Boutiques: a flexible framework for automated application integration in computing platforms}

@@

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, \\
  Pierre Bellec, Remi Bernard, Sorina Camarasu-Pop, \\
  Fr\'ed\'eric Cervenansky, Rafael Ferreira da Silva, \\ 
  John Flavin, Pascal Girard, Krzysztof J. Gorgolewski, \\
  Nathaniel Kofalt, Pierre-Olivier Quirion, Pierre Rioux, \\
  Gunnar Schaeffer, Marc-\'Etienne Rousseau, Alan C. Evans}

@@

Why Linux. Define application. Define computing platforms. Explain the problem and why it is difficult.

@@

\begin{itemize}
\item JSON manifest
\item Link to container
\item Invocation schema
\end{itemize}

@@

The core component of the tool JSON manifest is a flexible
command-line description represented as a simple string complying to
the syntax of the \texttt{bash} Linux shell interpreter. The
command-line string may contain placeholders for input and output
values. The placeholders may be any string, and the command line may
actually encompass several commands, e.g., separated by semicolumns,
pipes (\texttt{|}) or ampersands (\texttt{\&}). Such a flexible
command-line specification is meant to facilitate minor adjustments on
the command-line, for instance input decompression and output
archival, avoiding the use of multiple ``shims'' to align mismatching
applications in a pipeline~\cite{hull2004treating}.

Here is an example of a typical command-line description:
\begin{verbatim}
exampleTool_1 [STRING_INPUT] [FILE_INPUT] [ENUM_INPUT] | \
exampleTool_2 [FLAG_INPUT] [NUMBER_INPUT] >> [LOG].txt
\end{verbatim}
The command line description contains six command-line keys for input
and output values. When the application is executed, such keys will be
replaced by numerical values and file names according to the user
input. No particular syntax is imposed for the command-line keys, they
just have to be unique. Flags will also be added wherever
appropriate. Note the use of the pipe operator to chain tools, and of
the \texttt{>>} bash operator that allows to redirect the standard
output in a particular file.

\subsection{Input description}

\paragraph{General properties.} Inputs must have a name, a unique
identifier and a type. They may be optional, have a description, a
command-line key (placeholder), a flag and flag separator, and a
default value. Inputs may also be lists.

\paragraph{Types.} Inputs may be of type \texttt{String},
\texttt{Number}, \texttt{Flag} or \texttt{File}. Types can be
restricted to a specific set of values (\texttt{String} and
\texttt{Number} only), to a specific range (\texttt{Number} only) to
integers or booleans (\texttt{Number} only).

\paragraph{Groups and dependencies.} Groups of inputs may be defined
from a identifier, name and list of input identifiers. Groups may be
used for presentation purposes in a graphical user interface and to
specify the following constraints among inputs: (1)
\texttt{mutually-exclusive}: only one input in the group may have a
value (2) \texttt{one-is-required}: at least one of the inputs in the
group must have a value. Dependencies among inputs may also be defined
regardless of a particular group: a specific input may (1) require a
list of other inputs to be active to be available and (2) define a
list of inputs that are disabled when it is active.

Listing~\ref{listing:input-example} shows the definition of an input
in the command line exemplified above. According to this definition
and assuming that the input value entered by the user is 0.3, the
string \texttt{[NUMBER\_INPUT]} will be replaced by \texttt{-n=0.3} on
the command line at execution time.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    ``id'' : ``num_input'',
    ``name'' : ``A number input'',
    ``type'' : ``Number'',
    ``command-line-key'' : ``[NUMBER_INPUT]'',
    ``optional'' : true,
    ``command-line-flag'' : ``-n'',
    ``command-line-separator'' : ``='',
    ``minimum'' : 0,
    ``maximum'' : 1,
    ``exclusive-minimum'' : true,
    ``exclusive-maximum'' : false
  }
\end{minted}
\caption{Input example} 
\label{listing:input-example}
\end{listing}

\subsection{Output description}

@@

\subsection{Workflow support}

@@

The implementation of the application is defined in a container image
in the Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for
application developers and users who want to use applications on their
local workstation. It is very well documented, maintained and it has a
rich ecosystem of tools to help build and run containers on most
platforms, including Windows, MacOS and Linux. Singularity is more
suited for users and platforms that need to run applications on shared
computing clusters. Bridges exist among these containers formats to
convert container images across frameworks. For instance, a platform
dedicated to high-performance computing may accept manifests referring
to Docker containers to facilitate application integration by
developers, and internally convert container images to Singularity to
run applications efficiently on clusters.

Container images are defined from their URL (Singularity and rootfs)
or name in a Docker index. Container images may specify a working
directory where the application has to be run. A hash may also be
reported to accurately identify container images and detect updates.

\subsection{Custom properties}

\section{Implementation and Evaluation}

\subsection{Core tools} Validator, local executor, invocation schema generator.

\subsection{Platforms}

\paragraph{CBRAIN}

\paragraph{Pegasus}

\paragraph{VIP}

@@

@@

A table summarizing which Boutiques features are used in which pipelines.

@@

\section{Acknowledgments}

Amazon grant.

\section{References}
\bibliographystyle{plain}
\bibliography{biblio}
